### 2015-06-05 intro.tex
modified the introduction

%!TEX root =  autocontgrlp.tex
\section{Introduction}\label{intro}
Optimal sequential decision making problems occurring in science, engineering and economics can be cast in the framework of Markov Decision Processes (MDPs). A policy $u$, is a mapping from state space to action space and it specifies the action selection mechanism. The problem of \emph{control} requires coming up with good policies and for which one needs to \emph{predict} the behavior of individual policies. The problem of \emph{prediction} deals with computing the value-function $J_u$(a mapping from state space to reals\footnote{Also known as reward/cost-to-go function.}) which specifies the value of each state under a given policy $u$. In order to solve an MDP it is necessary to solve the problem of prediction as well as the problem of control.\todoc{Do we have a source to cite here? What does it mean to solve an MDP anyways? Find an optimal policy? Find a near-optimal policy? I think the story is that knowing the optimal value function helps, but I have not seen any claim that it is \emph{necessary} to compute it to find.}

Given an MDP,\todoc{How is the MDP given? If it is given as a list of values, just reading the values may take too much time.
Or we assume sparsity? Or we assume that one can query the next state distribution at any state-action? These yield to different models. We should specify which one we work with (presumably the sparse is the easiest).
} ideally, we would be interested in computing the optimal value function $J^*$ and the optimal policy $u^*$. The Bellman equation (BE) relates these quantities. Conventional solution methods for MDPs \cite{BertB} such as value iteration, policy iteration and linear programming can compute exact values of $J^*$ and $u^*$ by either directly or indirectly solving the Bellman equation. By computing $J^*$ and $u^*$, the conventional methods address both the prediction and control problems. However, in most practical problems the size of the state space grows exponentially in the number of state variables, making the computation of $u^*$ and $J^*$ intractable using the exact methods.

\emph{Approximate dynamic programming} (ADP) \cite{dpchapter,powell} refers to a gamut of approximate solution methods to MDPs with large number of states. In order to reduce the computational effort, ADP methods resort to dimensionality reduction by parameterizing the value function and/or the policy and restricting their search space to the chosen parameterized class. Thus, instead of computing exact values of $u^*$ and $J^*$, ADP methods only compute an approximate value function $\tilde{J}$ and a sub-optimal policy $\tilde{u}$. The quality of the ADP method is measured by the error $||J^*-\tilde{J}||$, in approximating the value function and the loss in performance due to the sub-optimal policy $||J^*-J_{\tilde{u}}||$, measured with respect to the optimal policy. Of the two error terms, $||J^*-J_{\tu}||$ is more important if we are interested in the problem of finding a useful policy. In the context of ADP, the prediction and control problems are said to be addressed when the errors $||J^*-\tilde{J}||$ and $||J^*-J_{\tilde{u}}||$ can be bounded so that the loss in performance due to the approximation can be quantified.
\todoc{I don't understand why one should care about  approximating $J^*$ at all.}

Value-function-only\footnote{Since this paper is concerned with results related to the approximate linear programming formulation we do not discuss other classes of ADP methods such as \emph{actor-only} and \emph{actor-critic} methods.} (or \emph{critic-only}) ADP methods approximate the value function and use it to obtain a one-step look ahead policy by making use of the Bellman equation. Most widely used parameterization to approximate the value function is the \emph{linear function approximator} (LFA). Methods employing LFA approximate the value function as $J_u\approx\Phi r^*$, where $\Phi$ is a feature matrix whose columns are the basis functions and $r^*$ is a weight vector to be learnt. Dimensionality reduction (in terms of the number of unknowns) is achieved by choosing fewer number of basis functions compared to the number of states.

A host of value-function-based ADP methods solve the \emph{projected Bellman equation} (PBE) which is obtained by including the least squares projection operator in the BE so as to project quantities onto the lower dimensional subspace. A limitation of using the least squares projection operator is that it reduces the prediction error $L_2$-norm and cannot ensure policy improvement which requires prediction error to be minimum in the $L_\infty$-norm. Due to this norm mismatch, policy improvement cannot be guaranteed and \emph{policy-chattering} (policy-oscillations) can occur in practice \cite{dpchapter}. 
As a result, the performance loss is difficult to control and existing result that control this loss are proven under restrictive conditions \cite[e.g.,][]{anszemu:mlj07,FaMuSz10}.

The \emph{approximate linear programming} (ALP) \cite{ALP,CS,SALP,ALP-Bor,gkp,fs,npalp} formulation also employs LFA, however, solves a linear program to compute $J^*\approx\tilde{J}=\Phi r^*$, and computes a one-step greedy policy. Since ALP computes an approximation to $J^*$ and outputs only a single policy $\tilde{u}$, there is no issue of policy-chattering. The ALP formulation also provides theoretical performance guarantees, i.e., both error terms $||J^*-\tilde{J}||$ and $||J^*-J_{\tilde{u}}||$ can be bounded. This makes ALP an attractive method since it addresses both the prediction and control problems. However, an important shortcoming of ALP is that the number of constraints are of the order of the state space, making the vanilla version of ALP intractable.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Introduction}\label{intro}
Optimal sequential decision making problems occurring in science, engineering and economics can be cast in the framework of Markov Decision Processes (MDPs). A policy $u$, is a mapping from state space to action space and it specifies the action selection mechanism. The problem of \emph{control} requires coming up with good policies and for which one needs to \emph{predict} the behavior of individual policies. The problem of \emph{prediction} deals with computing the value-function $J_u$(a mapping from state space to reals\footnote{Also known as reward/cost-to-go function.}) which specifies the value of each state under a given policy $u$. \todoc{How is the MDP given? If it is given as a list of values, just reading the values may take too much time.
Or we assume sparsity? Or we assume that one can query the next state distribution at any state-action? These yield to different models. We should specify which one we work with (presumably the sparse is the easiest).
}Ideally, we are be interested in computing the optimal value function $J^*$ and the optimal policy $u^*$. The Bellman equation (BE) relates these quantities and in particular $u^*$ can be obtained as a `greedy/one-step look ahead' policy by substituting $J^*$ in the BE. Conventional solution methods for MDPs \cite{BertB} such as value iteration, policy iteration and linear programming can compute exact values of $J^*$ and $u^*$ by either directly or indirectly solving the Bellman equation. By computing $J^*$ and $u^*$, the conventional methods address both the prediction and control problems. However, in the case of MDPs with large number of states, computing $u^*$ and $J^*$ using the exact methods is infeasible. \todoc{Do we have a source to cite here? What does it mean to solve an MDP anyways? Find an optimal policy? Find a near-optimal policy? I think the story is that knowing the optimal value function helps, but I have not seen any claim that it is \emph{necessary} to compute it to find.}

\emph{Approximate dynamic programming} (ADP) \cite{dpchapter,powell} refers to a gamut of approximate solution methods to MDPs with large number of states. Value-function-only\footnote{Since this paper is concerned with results related to the approximate linear programming formulation we do not discuss other classes of ADP methods such as \emph{actor-only} and \emph{actor-critic} methods.} (or \emph{critic-only}) ADP methods compute an approximate value function and use it to obtain a one-step look ahead policy by making use of BE. In order to reduce the computational effort, the approximate value function is restricted to a chosen parametrized class of functions. The most widely used parametrization to approximate the value function is the \emph{linear function approximator} (LFA). Methods employing LFA approximate the value function as $\Phi r^*$, where $\Phi$ is a feature matrix whose columns are the basis functions and $r^*$ is a weight vector to be learned. Computational advantage (in terms of the number of unknowns) is achieved by choosing fewer number of basis functions compared to the number of states. Once the set of basis function is chosen, the way to compute $r^*$ varies across the different ADP methods.\todoc{I don't understand why one should care about  approximating $J^*$ at all.}

A host of value-function-based ADP methods solve the \emph{projected Bellman equation} (PBE) which is obtained by including the least squares projection operator in the BE so as to project quantities onto the lower dimensional subspace. A limitation of using the least squares projection operator is that it reduces the prediction error $L_2$-norm and cannot ensure policy improvement which requires prediction error to be minimum in the $L_\infty$-norm. Due to this norm mismatch, policy improvement cannot be guaranteed and \emph{policy-chattering} (policy-oscillations) can occur in practice \cite{dpchapter}. As a result, the performance loss is difficult to control and existing result that control this loss are proven under restrictive conditions \cite{anszemu:mlj07,FaMuSz10}.

The \emph{approximate linear programming} (ALP) \cite{ALP,CS,SALP,ALP-Bor,gkp,fs,npalp} formulation also employs LFA, however, solves a linear program to compute $J^*\approx\tilde{J}=\Phi r^*$, and computes a one-step greedy policy $\tilde{u}$. Since ALP computes an approximation to $J^*$ and outputs only a single policy $\tilde{u}$, there is no issue of policy-chattering. The ALP formulation also provides theoretical performance guarantees, i.e., the prediction error $||J^*-\tilde{J}||$ and the control error\footnote{loss in performance due to the sub-optimal policy} $||J^*-J_{\tilde{u}}||$ can be bounded. This makes ALP an attractive method since it addresses both the prediction and control problems by providing the error bounds. However, an important shortcoming of ALP is that the number of constraints are of the order of the state space, making the vanilla version of ALP intractable.
